\bfseries How do sailboats work? A simple momentum-transport model

### 2024-08-21 paper/main.tex
tiny things
a few notes
caption tweak

% people who might comment on it:
% -------------------------------
% - KSF
% - Kilian Walsh
% - Sanjoy Mahajan
% - David Grier
% - Dani Maoz? Joe Shields?
% - All those in the acks.

@@

We (the authors) are not serious sailors.
However, we believe that sailing is a beautiful technology.
The boat is powered by the wind, and yet it can travel (possibly with a tacking course) in any direction.
Of course it's not consistent with Galilean relativity\footnote{We refer to Galilean relativity repeatedly in this \documentname.
Of course Galilean relativity is just the low-velocity approximation to special relativity \cite{sr}.
But we are not going to consider air--water velocity differences that are an appreciable fraction of the speed of light (except maybe in the discussion in \secref{sec:discussion}).} to say that the boat is powered by the wind:
There is no absolute velocity, and there is a valid rest frame in which the wind velocity is zero.
The Galilean-relativistic thing to say is that the boat is powered by the relative velocity between the air and the water.
Indeed, the symmetry between the air and the water in the mechanisms of sailing is the fundamental theme of one of the better books on the physics of sailing \cite{symmetry}.
Here we build the simplest possible physical model for a sailboat---and sailing---that is consistent with symmetries and captures all of the leading-order features of the technology.

@@

@@

@@

As we change the ratios of the effective areas, the shapes of the velocity curves change.
In \figref{fig:design} we show the effect on the boat velocity curves of changing the sail-to-keel ratio $A_\sail/A_\keel$, the sail lift ratio $A_\sail/A_\above$, and the keel lift ratio $A_\keel/A_\below$ (the keel lift ratio).
There are a number of boat-design considerations visible in this \figurename:
Sailing upwind is possible only if either the sail lift ratio is large, or the sail-to-keel ratio is small, or both.
Since upwind sailing is critical for the success of a boat, and since obtaining very high lift ratios can be difficult, we predict that most commercial sailboats have sail-to-keel ratios smaller than the 700 of our fiducial boat.
Sailing downwind faster than the wind is possible only if the keel lift ratio is large, or the sail-to-keel ratio is large.
Sailing cross-wind is at its best when the sail-to-keel ratio is equal to the ratio of the density of water to the density of air.
For context, recall that these ratios (lift ratios and sail-to-keel ratio) are ratios of effective areas, not real areas.
\begin{figure}[t!]
  \includegraphics[width=0.5\textwidth]{design_A.pdf}%
  \includegraphics[width=0.5\textwidth]{design_C.pdf}\\%
  \rule{0.25\textwidth}{0pt}\includegraphics[width=0.5\textwidth]{design_E.pdf}
  \caption{Each panel is analagous to \figref{fig:hourgood} but in each case we have varied one of the boat's dimensionless ratios. In each panel, the grey points show the same speed curve as that plotted in \figref{fig:hourgood}, which was computed for the boat described by the parameters given in \eqref{eq:boat1} through \eqref{eq:boat4}, and for a wind speed (relative to the water) of $8\,\mps$.
  In the top-left panel, the small and large black points show variations as we vary the sail lift ratio $A_\sail/A_\above$ from 40 (small dots) to 200 (grey dots) to 1000 (large dots).
  In the top-right panel, the small and large black dots show variations as we vary the keel lift ratio $A_\keel/A_\below$ from 40 (small dots) to 200 (grey dots) to 1000 (large dots).
  In the bottom panel, the small and large black dots show the variations as we vary the sail-to-keel ratio $A_\sail/A_\keel$ from 140 (small dots) to 700 (grey dots) to 3500 (large dots).\label{fig:design}}
  \figurerule
\end{figure}

@@

KLEBAN SAY: Should we quantitatively report on the conditions for dwfttw? And the conditions for upwind? Right now this is all qualitative.
HOGG SAY: YES; THERE are settings for the dimensionless ratios that permit and don't permit it. Not sure how to present this because there are three dimensionless ratios. Working on it.

>>>>

% people who might comment on it:
% -------------------------------
% - KSF
% - Kilian Walsh
% - Sanjoy Mahajan
% - David Grier
% - Dani Maoz? Joe Shields?
% - Saskia Hekker
% - All those in the acks.

@@

We (the authors) are not serious sailors!
However, we believe that sailing is a beautiful technology.
The boat is powered by the wind, and yet it can travel (possibly with a tacking course) in any direction.
Of course it's not consistent with Galilean relativity\footnote{We refer to Galilean relativity repeatedly in this \documentname.
Of course Galilean relativity is just the low-velocity approximation to special relativity \cite{sr}.
But we are not going to consider air--water velocity differences that are an appreciable fraction of the speed of light (except maybe in the discussion in \secref{sec:discussion}).} to say that the boat is powered by the wind:
There is no absolute velocity, and there is a valid rest frame in which the wind velocity is zero.
The Galilean-relativistic thing to say is that the boat is powered by the relative velocity between the air and the water.
Indeed, the symmetry between the air and the water in the mechanisms of sailing is the fundamental theme of one of the better books on the physics of sailing \cite{symmetry}.
Here we build the simplest possible physical model for a sailboat---and sailing---that is consistent with symmetries and captures all of the leading-order features of the technology.

@@

HOGG: How are uw and dwfttw the same question in different reference frames?

@@

HOGG: Note about how it feels. Is this repetitive with comments above? If so, remove them and replace with references down to this Section.

@@

As we change the ratios of the effective areas, the shapes of the velocity curves change.
In \figref{fig:design} we show the effect on the boat velocity curves of changing the sail-to-keel ratio $A_\sail/A_\keel$, the sail lift ratio $A_\sail/A_\above$, and the keel lift ratio $A_\keel/A_\below$ (the keel lift ratio).
There are a number of boat-design considerations visible in this \figurename:
Sailing upwind is possible only if either the sail lift ratio is large, or the sail-to-keel ratio is small, or both.
Since upwind sailing is critical for the success of a boat, and since obtaining very high lift ratios can be difficult, we predict that most commercial sailboats have sail-to-keel ratios smaller than the 700 of our fiducial boat.
Sailing downwind faster than the wind is possible only if the keel lift ratio is large, or the sail-to-keel ratio is large.
Sailing cross-wind is at its best when the sail-to-keel ratio is equal to the ratio of the density of water to the density of air.
For context, recall that these ratios (lift ratios and sail-to-keel ratio) are ratios of effective areas, not real areas.
\begin{figure}[t!]
  \includegraphics[width=0.5\textwidth]{design_A.pdf}%
  \includegraphics[width=0.5\textwidth]{design_C.pdf}\\%
  \rule{0.25\textwidth}{0pt}\includegraphics[width=0.5\textwidth]{design_E.pdf}
  \caption{HOGG SOME PARTS OF THIS CAPTION ARE WRONG Each panel is analagous to \figref{fig:hourgood} but in each case we have varied one of the boat's dimensionless ratios. In each panel, the grey points show the same speed curve as that plotted in \figref{fig:hourgood}, which was computed for the boat described by the parameters given in \eqref{eq:boat1} through \eqref{eq:boat4}, and for a wind speed (relative to the water) of $8\,\mps$.
  In the top-left panel, the small and large black points show variations as we vary the sail lift ratio $A_\sail/A_\above$ from 40 (small dots) to 200 (grey dots) to 1000 (large dots).
  In the top-right panel, the small and large black dots show variations as we vary the keel lift ratio $A_\keel/A_\below$ from 40 (small dots) to 200 (grey dots) to 1000 (large dots).
  In the bottom panel, the small and large black dots show the variations as we vary the sail-to-keel ratio $A_\sail/A_\keel$ from 140 (small dots) to 700 (grey dots) to 3500 (large dots).\label{fig:design}}
  \figurerule
\end{figure}

KLEBAN SAY: Should we quantitatively report on the conditions for dwfttw? And the conditions for upwind? Right now this is all qualitative.
HOGG SAY: YES; THERE are settings for the dimensionless ratios that permit and don't permit it. These are illustrated in \figref{fig:dwfttw}, as a function of lift ratios, for boats with two different values of the sail-to-keel area ratio.
The brief summary is .... HOGG
\begin{figure}[t!]
  \includegraphics[width=0.5\textwidth]{upwind_vmg.pdf}%
  \includegraphics[width=0.5\textwidth]{downwind_vmg.pdf}\\%
  \rule{0.25\textwidth}{0pt}\includegraphics[width=0.5\textwidth]{crosswind_vmg.pdf}
  \caption{HOGG WRITE CAPTION.
  In the top-left panel, HOGG.
  In the top-right panel, HOGG.
  In the bottom panel, HOGG.\label{fig:dwfttw}}
  \figurerule
\end{figure}

@@
